---
title: "Code Snippets"
execute: 
  eval: false
code-block-bg: "white"
code-block-border-left: "black"
---

# Code Snippets

Morgan-Wall T (2025). *rayshader: Create Maps and Visualize Data in 2D and 3D*. R package version 0.38.10, https://github.com/tylermorganwall/rayshader, [https://www.rayshader.com](https://www.rayshader.com/).

Roussel J, Auty D, Coops NC, Tompalski P, Goodbody TR, Meador AS, Bourdon J, de Boissieu F, Achim A (2020). “lidR: An R package for analysis of Airborne Laser Scanning (ALS) data.” *Remote Sensing of Environment*, **251**, 112061. ISSN 0034-4257, [doi:10.1016/j.rse.2020.112061](https://doi.org/10.1016/j.rse.2020.112061), <https://www.sciencedirect.com/science/article/pii/S0034425720304314>.

Roussel J, Auty D (2024). *Airborne LiDAR Data Manipulation and Visualization for Forestry Applications*. R package version 4.1.2, <https://cran.r-project.org/package=lidR>.

### Shade simulation function in R

```{r, eval=FALSE}
#library(suncalc)
#library(rayshader)
#library(terra)

#### Setup time intervals and sun positions

#start <-  as.POSIXct("2024-07-17 01:00:00", tz = "America/Los_Angeles")
#end <-  as.POSIXct("2024-07-18 00:00:00", tz = "America/Los_Angeles")

#times <- data.frame(dates = seq(start, end, "hour"))

#sun_pos <- getSunlightPosition(date = times$dates, lat = 49.2827, lon = -123.1207)

shade_sim <- function(pc, sun_calc,name){   #take a LiDAR point cloud,
  alt <- sun_calc$altitude*(180/pi)         #a table of sun positions,
  azim <- sun_calc$azimuth*(180/pi)         #and output file name as inputs
  
  if(alt >= 0){
    dsm <- rasterize_canopy(pc, dsmtin(), res = 2) %>%    #create and shade a dsm
      raster_to_matrix() %>% 
      ray_shade(sunaltitude = alt, sunangle = azim) %>% 
      rast()
  
    writeRaster(dsm, name, overwrite = TRUE)   #write out the result
  }
}

```

### Isolate center tree from a LiDAR point cloud

```{r, eval = FALSE}
#Load in 764 segmented point clouds
path <- (list.files("Outputs/Species Level/las_clipped", full.names = TRUE))

rois <- list()
for(i in 1:length(path)){
  rois[[i]] <- readLAS(path[i])
}

#Point clouds are circular regions clipped from larger file. This allowed for the center coordinates to be included in the filename and used in isolating the central tree below
roi_indiv <- list()
for(i in 1:764){
#generate metrics as a table of treeIDs and coordinates
metrics <- crown_metrics(rois[[i]], ~list(z_max = max(Z), z_mean = mean(Z)))

#use filename to create point geometries in the center of the plot
test_centroid <- str_extract(path[i], 
                 "[:digit:]{6}\\.[:digit:]{1}\\_[:digit:]{7}|[:digit:]{6}\\_[:digit:]{7}") %>%
  str_replace("_", ", ") %>%
  str_split("\\,\\s",2) %>%
  unlist() %>%
  as.numeric() %>%
  st_point() %>%
  st_geometry()

st_crs(test_centroid) <- 26910
  
#add new row to metrics table that contains distance from tree to center
new_metrics <- mutate(metrics, dist = st_distance(metrics, test_centroid))
  
#central tree is the tree with the shortest distance to the centroid
center_tree <- filter(new_metrics, dist == min(dist))
  
#assign it back out to a list
roi_indiv[[i]] <- filter_poi(rois[[i]], treeID == center_tree$treeID)
}
```

### Arcpy Script for Shade Percentage Analysis

``` python
#Loop georeference function through each folder, georeference shade simulation outputs
og_path = r"C:\Users\mlefcoe.stu\Documents\Shade_Mapping_R\Outputs\Neighbourhood Level"
control_path = r"C:\Users\mlefcoe.stu\Documents\Shade_Mapping_R\Inputs\Control_Points"
ras_folders = os.listdir(og_path)

for i in range(len(ras_folders)):
    georefer(directory = og_path + "\\" + ras_folders[i], out_name =
    ras_folders[i][:ras_folders[i].index(" ")], link = control_path + "\\" +
    ras_folders[i][:ras_folders[i].index("_")] + ".txt")


#Define function to reclassify all shade rasters in a folder
def slicer(directory, out_name):
    count = 0
    ras_files = os.listdir(directory)
    for i in range(len(ras_files)):
        if ras_files[i].endswith(".tif"):
            ras = directory + "\\" + ras_files[i]
            out_slice = arcpy.sa.Slice(in_raster=ras, number_zones=3, 
            slice_type = "NATURAL_BREAKS")
            out_slice.save(directory + "\\" + out_name + "reclassed" + 
            str(count) + ".tif")
            count = count + 1

#Loop reclassing functions through folders
og_path = r"C:\Users\mlefcoe.stu\Documents\Shade_Mapping_R\Outputs\Neighbourhood Level"
ras_folders = os.listdir(og_path)

for i in range(len(ras_folders)):
    slicer(directory = og_path + "\\" + ras_folders[i], 
    out_name=ras_folders[i][:ras_folders[i].index(" ")])

#Define function to extract shade from all shade rasters in a folder
def extract_shade(directory, out_name):
    count = 0
    ras_files = os.listdir(directory)
    for i in range(len(ras_files)):
        if ras_files[i].endswith("reclassed" + str(count) + ".tif"):
            ras = directory + "\\" + ras_files[i]
            attextract = arcpy.sa.ExtractByAttributes(in_raster = ras,
            where_clause = "VALUE = 1")
            attextract.save(directory + "\\" + out_name + "extracted" + 
            str(count) + ".tif")
            count = count + 1
    count = 10
    for i in range(len(ras_files)):
        if ras_files[i].endswith("reclassed" + str(count) + ".tif"):
            ras = directory + "\\" + ras_files[i]
            attextract = arcpy.sa.ExtractByAttributes(in_raster = ras,
            where_clause = "VALUE = 1")
            attextract.save(directory + "\\" + out_name + "extracted" + 
            str(count) + ".tif")
            count = count + 1

#Loop extracting function through each folder
og_path = r"C:\Users\mlefcoe.stu\Documents\Shade_Mapping_R\Outputs\Neighbourhood Level"
ras_folders = os.listdir(og_path)

for i in range(len(ras_folders)):
    extract_shade(directory = og_path + "\\" + ras_folders[i], 
    out_name=ras_folders[i][:ras_folders[i].index(" ")])

#Define a function to convert shade rasters in a folder to polygons
def ras_to_poly(directory, out_name):
    count = 0
    ras_files = os.listdir(directory)
    for i in range(len(ras_files)):
        if ras_files[i].endswith("extracted" + str(count) + ".tif"):
            ras = directory + "\\" + ras_files[i]
            arcpy.conversion.RasterToPolygon(in_raster=ras, 
            out_polygon_features = directory + "\\" + out_name + "shade_poly" +
            str(count) + ".shp")
            count = count + 1
    count = 10
    ras_files = os.listdir(directory)
    for i in range(len(ras_files)):
        if ras_files[i].endswith("extracted" + str(count) + ".tif"):
            ras = directory + "\\" + ras_files[i]
            arcpy.conversion.RasterToPolygon(in_raster=ras, 
            out_polygon_features = directory + "\\" + out_name + "shade_poly" +
            str(count) + ".shp")
            count = count + 1

#Loop polyshade function through all folders
og_path = r"C:\Users\mlefcoe.stu\Documents\Shade_Mapping_R\Outputs\Neighbourhood Level"
ras_folders = os.listdir(og_path)
ras_folders

for i in range(len(ras_folders)):
    ras_to_poly(directory = og_path + "\\" + ras_folders[i], 
    out_name=ras_folders[i][:ras_folders[i].index(" ")])

#Define a function to interesect each shade raster in a folder with a specified feature class
def intersector(directory, out_name, feature):
    count = 0
    ras_files = os.listdir(directory)
    for i in range(len(ras_files)):
        if ras_files[i].endswith("poly" + str(count) + ".shp"):
            ras = directory + "\\" + ras_files[i]
            arcpy.analysis.PairwiseIntersect(in_features = [feature, ras],
            out_feature_class = directory + "\\" + out_name + "intersected" +
            str(count))
            count = count + 1
    count = 10
    for i in range(len(ras_files)):
        if ras_files[i].endswith("poly" + str(count) + ".shp"):
            ras = directory + "\\" + ras_files[i]
            arcpy.analysis.PairwiseIntersect(in_features = [feature, ras],
            out_feature_class = directory + "\\" + out_name + "intersected" +
            str(count))
            count = count + 1

#Loop the intersection function through each folder
for i in range(len(ras_folders)):
    intersector(directory = og_path + "\\" + ras_folders[i], 
    out_name = ras_folders[i][:ras_folders[i].index(" ")], 
    feature = r"C:\Users\mlefcoe.stu\OneDrive - UBC\Documents\FCOR599\Shade         Mapping\building_polygons.shp")
```
